---
title: "Spatial-Temporal Distribution of Kinkajous in a Balsa Tree"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

# Overview

The analysis in this document seeks to answer the question "What factors influence the movements of kinkajous within the crown of a balsa tree?". We start with a few hypotheses:

A. Kinkajous prefer to forage in regions with greater densities of flowers
B. Kinkajous assess the density of flowers within a region through perception and/or sampling effort.
C. Kinkajous predict which regions will have more flowers using information obtained from foraging on previous nights
D. Kinkajous will avoid direct foraging competition with other kinkajous
E. Kinkajous remember where they have been and avoid foraging in recently depleted regions

To test these hypotheses, we collected data on the location of flowers within a balsa tree crown each night, and the locations of kinkajous within that crown every minute for four-hour periods. In order to conduct a spatial analysis, we divided the tree crown into three regions. As best as possible, these regions divided the crown such that all the offshoots of one or more primary branches were contained within a single region. We then tested how the spatio-temporal distribution of flowers within these regions related to the residence times of kinkajous within each region, and to the transitions of kinkajous between regions.

In this document, we 1) prepare and clean the data, 2) Develop predictions about the relationships between flower counts, kinkajou presence, and kinajous movements within and between regions, 3) visualize these relationships, and 4) Develop and run statistical models to test predictions that appear to be supported by patterns in the data.

# Data Setup

```{r, setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(sf)
library(dplyr)
library(ggplot2)
library(scales)
library(nlme)
```

```{r helper_functions, include=FALSE}
get_study_day_and_time <- function(Datetime_string, study_start) {
  #study start must include the approximate start time of evening observations if overnight follows are to be binned into the correct study day
  datetime_vector <- strsplit(Datetime_string, split = "")[[1]]
  date1 <- paste(paste(datetime_vector[1:6], collapse = ""), 
                "20", 
                paste(datetime_vector[7:8], collapse = ""), sep = "") %>% as.POSIXct(format = "%d-%m-%Y", tz = "America/Panama")
  time1 = paste(datetime_vector[9:10], collapse = "")
  study_day = difftime(date1, study_start, units = "days") %>% floor()
  return(list(study_day, time1))
}

get_kinkMinutesPerHour <- function(count_data, categories = 6) {
  #default value for categories assumes count data has 6 rows per observational minute, representing 3 Regions x 2 states (hidden or visible)
  minutes_per_hour = (sum(count_data)/(length(count_data)/categories)) * 60
  return(minutes_per_hour)
}

get_growth_index <- function(time_series) {
  removed = 0
  while(time_series[1] == 0) {
    time_series <- time_series[-1]
    warning("initial observation removed due to value of 0")
    removed = removed + 1
  }
  growth_index <- ((time_series-time_series[1])/time_series[1]) * 100
  return(c(rep(NA, times = removed), growth_index))
}
```

The data used for this analysis come from two sources: side-angle thermal video recordings of the balsa tree and top-down photographs of the tree captured by drone.

## Thermal Video Data Processing and Kinkajou Counts

The thermal videos are of the full balsa tree crown, recorded from the side. Most recordings were four hours, except where interrupted by rain. Recordings had outlines of the regions superimposed by Alexander Vining (using the software ProCreate), who observed the balsa tree in person during all recordings. Because these files are large, they are not included in this repository. Instead, the data included here are the number of kinkajous in each region for each minute of a recording. These data are contained in the folder Kinkajou_Scans_Nele. The counts were determined by Nele Stockmeyer, who watched all of the thermal videos and scanned the tree each minute. Kinkajou counts are further annoted by the number of counted kinkajous that were not visible during the given scan, but their presence was inferred from other scans ("hidden").  Further details on the data collection process are contained in the manuscript ". . ."

In addition to the kinkajou counts, which are annotated by video-minutes, the file video_file_timestamp_metdata_cam1 contains the start and end timestamps of each video. These allow conversion of video-minutes to timestamps. 

We begin data preparation by determining the file names for all of the kinkajous count scan data, for which there is one .csv file per night of recording. We also parse the metadata file to store the start timestamps in a vector. We then read in each kinkajous count scan data file, convert the video-minutes to timestamps, and collect all of the data into a single dataframe.

```{r data_load, message = FALSE}
file_names <- dir("../DATA/raw/Kinkajou_Scans_Nele", pattern = ".csv")

video_metadata_location <- "../DATA/raw/Video_metadata/video_file_timestamp_metadata_cam1.txt"
video_metadata <- readChar(con = video_metadata_location, nchars = file.info(video_metadata_location)$size)
video_metadata <- str_split(video_metadata, pattern = "Folder: ")[[1]][-1] #break all metadata into strings for each observation period (and remove preceding empty character)
video_metadata <- str_split(video_metadata, pattern =  "start = ") #find the start time of all files
start_timestamps <- sapply(video_metadata, function(X) str_split(X[2], pattern = "\r")[[1]][1]) #extract start time of first file and remove all extra data
start_timestamps <- as.POSIXct(start_timestamps, format = "%Y-%m-%d_%H-%M-%S", tz = "GMT")
start_timestamps <- with_tz(start_timestamps, "America/Panama")


all_kinkajou_scans <-  c()
Jan_22_check = FALSE #used later to parse two video files that started on the same day (one AM one PM)

for(file_name in file_names) {
  scan_data <- read_csv(paste("../DATA/raw/Kinkajou_Scans_Nele/", file_name, sep = ""), skip = 1, progress = FALSE)
  colnames(scan_data) <- c("timestamp", "Region1_total", "Region1_hidden", "Region2_total", "Region2_hidden", "Region3_total", "Region3_hidden", "Kinkajou_Notes", "Video_Notes", "Extraneous")
  scan_data <- scan_data[complete.cases(scan_data$timestamp),] %>% select(! Extraneous)
  start_date <- str_split(file_name ,pattern = "_Cam")[[1]][1] #str_split returns a list of string components. We want the first (and only) element of this list - then the first string element in this list gives the date
  start_date <- as.POSIXct(paste("20", start_date, sep = ""), format = "%Y_%m_%d") #date needs all 4 digits
  start_timestamp <- start_timestamps[which(date(start_timestamps) == start_date)]
  #Two video files started on Jan 22nd, one AM and one PM. Because this is the ONLY date where this occurred, I implemented a quick and dirty check to parse the AM file first and the PM file second, below
  if (length(start_timestamp) > 1)
      if (! Jan_22_check) {
        start_timestamp <- start_timestamp[1]
        Jan_22_check <-  TRUE
      } else start_timestamp <- start_timestamp[2]
  #end Jan 22 case
  scan_data$timestamp <- start_timestamp + period_to_seconds(hms(scan_data$timestamp))
  scan_data$Study_night <- round(as.numeric(start_timestamp - start_timestamps[1], units = "days"), digits = 0)
  all_kinkajou_scans <- rbind(all_kinkajou_scans, scan_data)
}
```

We continue preparing the kinkajous count data by converting the data from a) total kinkajous per region and b) hidden kinkajous per region to a) visible kinkajous per region and b) hidden kinkajous per region. This makes analysis using either type of count separately easier.

```{r data_prep}
#convert empty count cells (NAs) to 0
count_data <- all_kinkajou_scans[,2:7] # columns 2 through 7 are the count data columns, we want NAs in the notes columns, but 0s in the count data
count_data[is.na(count_data)] <- 0
all_kinkajou_scans[,2:7] <- count_data

#The data as read contain the TOTAL number of kinkajous and the number of those total that are hidden. As suggested by the new names given to the columns in this data structure, we want the count of visible and hidden kinkajous, not the total. The adjustment is made below
all_kinkajou_scans <- mutate(all_kinkajou_scans, Region1_visible = Region1_total - Region1_hidden, Region2_visible = Region2_total - Region2_hidden, Region3_visible = Region3_total - Region3_hidden) %>% 
  select(timestamp, Region1_total, Region2_total, Region3_total, Region1_visible, Region1_hidden, Region2_visible, Region2_hidden, Region3_visible, Region3_hidden, Study_night, Kinkajou_Notes, Video_Notes)

#pivot dataframe to long format for easy usage in ggplot
all_kinkajou_scans_long <- pivot_longer(all_kinkajou_scans,
                                   cols = Region1_visible:Region3_hidden,
                                   names_to = c("Region", "Visibility"),
                                   names_sep = "_",
                                   values_to = "Count")

all_kinkajou_scans_long$Region <- sapply(all_kinkajou_scans_long$Region, switch, "Region1" = "1", "Region2" = "2", "Region3" = "3") %>% factor()

write.csv(all_kinkajou_scans_long, "../DATA/processed/all_kinkajou_scans.csv")
```

In addition to the counts of kinkajous in each region, we are interested in how kinkajous transitions between region. Thus, we use the count data to create a transition matrix for each scan, where the number in each cell denotes how many kinkajous moved from the region denoted by the row to the region denoted by the column. These transitions matrices are collated into an array for each study night, where the third dimension is the scan (video-minute). The arrays for each night are stored in a list.

```{r Get_transitions}

get_transitions <- function(count_data){
  #input: a dataframe with columns "Region1_total", "Region2_total", "Region3_total" and (optionally) "timestamp".
  #output: an array giving the number of transitions from one region (first dimension) to another (second dimension) for each minute (row) of the input data (third dimension)
  transitions <- array(data = 0, dim = c(4,4,nrow(count_data)))
  if("timestamp" %in% colnames(count_data)) arrange(count_data, timestamp)
  count_data <- select(count_data, Region1_total, Region2_total, Region3_total)
  for(i in 2:nrow(count_data)){
    #to calculate the transition matrix, we treat the counts in the previous timestep as the row totals of the transition matrix
    row_totals <- c(unlist(count_data[i-1,]),
                    max(sum(count_data[i,]) - sum(count_data[i -1,]), 0)) #create a fourth row total for kinkajous entering the tree by differencing the total kinkajou counts of the current timestep and the previous timestep, counting only positive differences.
    #we then treat the counts in the current timestep as the column totals in the transition matrix
    col_totals <- c(unlist(count_data[i,]),
                    max(sum(count_data[i-1,]) - sum(count_data[i,]), 0)) #create a fourth column total for kinkajous leaving the tree
    for(j in seq_along(row_totals)){
      for(k in seq_along(col_totals)){
        m <- min(row_totals[j], col_totals[k]) #m is the maximum number of transitions that can contribute to both the relevant row total and column total
        transitions[j,k,i] <- m
        row_totals[j] <- row_totals[j] - m #subtract the number of transitions that have been added from the corresponding row total
        col_totals[k] <- col_totals[k] - m #subtract the number of transitions that have been added from the corresponding column total
      }
    }
  }
  return(transitions)
}

transition_arrays <- all_kinkajou_scans %>% group_by(Study_night) %>% group_map(~get_transitions(.x))
names(transition_arrays) <- as.character(unique(all_kinkajou_scans$Study_night))
```

Because 

## Drone Image Processing and Flower Counts

The second source of data are orthomosiac images of the balsa crown created by stitching together 100 images taken sequentially by a drone flying over the balsa tree. Because, again, these files are large, the data contained in this repository are geopackages containing WGS 84 locations and tree crown region of each flower in the crown. These data were extracted by first stitching together the drone photos from each flight into an orthmosaic (speak to Andreas about details). Second, Alexander Vining used the software QGIS to align each orthomosiac with Google satellite imagery of the study site and hand-delinieate the three regions of the tree crown. Finally, Nele Stockmeyer created spatial points for each flower in the orthomosaic, noting whether the flower appeared to upright or fallen over.

The following script extracts the spatial coordinates and regions from the geopackage files and stores them as an spatial features (sf) object, with additinal annotation for the date and time of the images from which data were collected.

```{r load_balsa_data, message = FALSE}
files = strsplit(dir("../DATA/raw/Flower_geopackages"), split = "[.]") #seperate filenames from extensions for all files in folder
files = files[sapply(files, function(X) X[2] == "gpkg")] #reduce file list only to gpkg files
geopackages <- vector("list", length = length(files))
for(i in seq_along(files)){
  flower_sf <- st_read(paste("../DATA/raw/Flower_geopackages/", files[[i]][1], ".", files[[i]][2], sep = ""), quiet = TRUE)
  flower_sf$Region <- strsplit(files[[i]][1], split = NULL)[[1]] %>% last() %>% as.numeric() + 1 #QGIS labelled regioning grouping outputs at 0 (last element of savefile string); add one to get region labels that match kinkajou data
  flower_sf$Date <- strsplit(files[[i]][1], split = "_")[[1]][2]
  geopackages[[i]] <- flower_sf
}
all_upflowers <- do.call(rbind, geopackages)
all_upflowers$Region <- factor(all_upflowers$Region)

#Drone flights are labelled by a character with date and time of scan. The date needs to be converted to a study day variable and the time put into its own column
all_upflowers <- all_upflowers %>% mutate(Study_night = sapply(Date, function(X) get_study_day_and_time(X, study_start = as.POSIXct("2019-12-14 18:00:00", tz = "America/Panama"))[[1]]),
                                          Time = factor(sapply(Date, function(X) get_study_day_and_time(X, study_start= as.POSIXct("2019-12-14 18:00:00", tz = "America/Panama")) [[2]])))

```

# Data Visualization

With the data prepared, we visualize spatio-temporal patterns of both flowers and kinkajous within the tree-crown as a whole and within regions of the tree crown. Based on our hypotheses, we made some predictions about patterns we expected to see


```{r data_visuals1}
plot_data <- ggplot(all_kinkajou_scans_long)

plot_data + 
  geom_bar(aes(x = Region, weight = Count)) + 
  facet_wrap(~Study_night) +
  theme_classic() +
  labs(y = "Kinkajous-Minutes")

ggplot(all_kinkajou_scans_long) +
  geom_bar(aes(x = Region, weight = Count, fill = Visibility)) +
  facet_wrap(~Study_night)
```



```{r viz_by_day}
ggplot(all_upflowers, aes(x = Study_night, color = Time)) + geom_line(stat = "count") + geom_point(stat = "count") +
  theme_classic() +
  labs(title = "Upright Flowers Counted")

ggplot(all_kinkajou_scans_long, aes(x = Study_night, y = Count)) + stat_summary(fun = "get_kinkMinutesPerHour", geom = "line") + stat_summary(fun = "get_kinkMinutesPerHour", geom = "point") +
  theme_classic() +
  labs(title = "Time Spent in Crown", y = "Kinkajou Minutes per Hour")

all_upflowers <- all_upflowers %>% filter(Time == "AM" | !Study_night %in% c(23, 26, 25, 32)) %>% #remove data from images marked in notes as poor quality photos
  filter(Time == "PM" | !Study_night %in% c(20,21)) #remove data from images marked in notes as poor quality photos


  
flower_counts <- all_upflowers %>% 
  group_by(Study_night, Time) %>% summarise(Count = n()) %>% 
  mutate(Study_night = Study_night - (Time == "AM")) %>% #study night is based on date, but we want to match AM recordings to previous PM recordins, so we subtract one from all AM study night values
  group_by(Study_night) %>% #where there were good photos for both AM and PM, we will use the mean flower count which is achieved by grouping and summariying
  summarise(Count = mean(Count)) %>% 
  transform(Growth = get_growth_index(Count)) #used to visually compare changes in flower count to changes in kinkajou residency without bias


kinkajou_counts <- all_kinkajou_scans_long %>% group_by(Study_night) %>% summarise(Minutes_Per_Hour = get_kinkMinutesPerHour(Count)) %>% transform(Growth = get_growth_index(Minutes_Per_Hour))
combined_data <- merge(flower_counts, kinkajou_counts, by = "Study_night", all = TRUE)

ggplot(flower_counts, aes(x = Study_night, y = Growth)) + 
  geom_line(stat = "identity") +
  geom_point(stat = "identity") +
  geom_line(data = kinkajou_counts, stat = "identity", linetype = "dashed") +
  geom_point(data = kinkajou_counts, stat = "identity") +
  theme_classic() +
  scale_color_discrete(name = "Flower Count") +
  labs(title = "Growth in Minutes per Hour Kinkajous spend in a balsa crown (black dotted) \n and balsa flower counts (solid) over study period",
       y = "Growth (%)",
       x = "Study Night")

ggplot(all_upflowers) + geom_sf(aes(color = factor(Date), shape = factor(Region)))
ggplot(all_upflowers) + geom_bar(aes(x = factor(Region))) + facet_wrap(~Date)

ggplot(flower_counts, aes(x = Study_night, y = Growth)) + 
  geom_line(aes(linetype = "Flower Count"), stat = "identity") +
  geom_point(stat = "identity") +
  geom_line(mapping = aes(linetype = "Residence Time"), data = kinkajou_counts, stat = "identity") +
  geom_point(data = kinkajou_counts, stat = "identity") +
  theme_classic() +
  scale_color_discrete(name = "Flower Count") +
  scale_linetype_manual(values = c("Flower Count" = "solid", "Residence Time" = "dashed")) +
  labs(#title = "Growth in Minutes per Hour Kinkajous spend in a balsa crown (black dotted) \n and balsa flower counts (solid) over study period",
       y = "Growth (%)",
       x = "Study Night")
```

```{r analysis_by_region}
flower_counts_region <- all_upflowers %>% mutate(Study_night = Study_night - (Time == "AM")) %>% 
  group_by(Study_night, Region) %>% 
  summarise(Count = n()) %>% 
  group_by(Study_night, Region) %>% 
  mutate(Count = mean(Count)) %>% 
  group_by(Region) %>% group_modify(~ transform(.x, Growth = get_growth_index(Count))) %>% 
  ungroup()

kinkajou_counts_region <- all_kinkajou_scans_long %>% group_by(Study_night, Region) %>% summarise(Minutes_Per_Hour = get_kinkMinutesPerHour(Count, categories = 2)) %>% group_by(Region) %>% group_modify(~ transform(.x, Growth = get_growth_index(Minutes_Per_Hour)))

ggplot(flower_counts_region, aes(x = Study_night, y = Count, color = Region)) + geom_line(stat = "identity") + geom_point(stat = "identity") +
  theme_classic() +
  labs(title = "Flower Counts by Region")

ggplot(kinkajou_counts_region, aes(x = Study_night, y = Minutes_Per_Hour, color = Region)) + geom_line(stat = "identity") + geom_point(stat = "identity") +
  theme_classic() +
  labs(title = "Time Spent by Kinkajous in Region")

ggplot(flower_counts_region, aes(x = Study_night, y = Growth)) + geom_line(color = "green3", stat = "identity") + geom_point(stat = "identity", color = "green3") +
  geom_line(data = kinkajou_counts_region, aes(color = "black")) + geom_point(data = kinkajou_counts_region, color = "black") +
  theme_classic() + 
  facet_wrap(~Region, labeller = "label_both") +
  theme(legend.position = "bottom") +
  scale_color_manual(name = "", values = c("Flower Count" = "green3", "Kinkajou Minutes Per Hour" = "black")) +
  labs(#title = "Growth in Minutes per Hour Kinkajous spend in a balsa crown (black) \n and balsa flower counts (solid) over study period",
       y = "Growth (%)",
       x = "Study Night")
```



```{r summary statistics}
#means and standard deviation of kinkajou movement
mean_residence <- mean(kinkajou_counts$Minutes_Per_Hour)
var_residence <- sd(kinkajou_counts$Minutes_Per_Hour)
mean_residence_by_region <- group_by(kinkajou_counts_region, Region) %>% dplyr::summarise(Mean = mean(Minutes_Per_Hour), Standard_Deviation =sd(Minutes_Per_Hour))
mean_residence_by_region$Region <- mean_residence_by_region$Region %>% as.character()
mean_residence_by_region <- mean_residence_by_region %>% rbind(c("all", mean_residence, var_residence))
mean_residence_by_region$Units <- "minutes per hour"
mean_residence_by_region

#means and standard deviation of balsa flower
mean_flowers_counts <- mean (flower_counts$Count)
var_flower_counts <- sd(flower_counts$Count)
mean_flower_counts_by_region <- group_by(flower_counts_region, Region) %>% summarise(Mean = mean (Count), Standard_Deviation = sd (Count))
mean_flower_counts_by_region$Region <- mean_flower_counts_by_region$Region %>% as.character()
mean_flower_counts_by_region <- mean_flower_counts_by_region %>% rbind(c("all", mean_flowers_counts, var_flower_counts))
mean_flower_counts_by_region$Units <- "Flowers"
mean_flower_counts_by_region

combined_summary_statistics <- rbind(mean_residence_by_region, mean_flower_counts_by_region)
combined_summary_statistics

paste("mean kinkajou residence in the entire tree, minutes per hour", mean_residence, 
      "standard deviation of kinkajou residence in entire tree, minutes per hour", var_residence,
      "mean flower count in entire tree by day", mean_flowers_counts,
      "standard deviation of mean flower count in entire tree by day", var_flower_counts, sep=" ")
```

```{r regression analysis}

#predicting flower counts based on study night
flowers_regression_all <- lm(data = select(as.data.frame(flower_counts), !geom), formula = Count ~ Study_night)
summary(flowers_regression_all)
#flowers_regression_all

plot(y = predict(flowers_regression_all, newdata = data.frame(Study_night = 0:max(flower_counts$Study_night))), x = 0:max(flower_counts$Study_night), type = "l",
     xlab = "Study Night",
     ylab = "Flower Count",
     main = "Full Tree Flower Count Regression")
points(x = flower_counts$Study_night, y = flower_counts$Count)

plot(x = flower_counts$Study_night, y = flowers_regression_all$residuals,
     xlab = "Study Night",
     ylab = "Residuals",
     main = "Full Tree Flower Count Residuals")
flowers_regression_all$effects

Sflower_all <- sqrt(sum(flowers_regression_all$residuals ^ 2)/(length(flowers_regression_all$residuals) - 2))
qqnorm(flowers_regression_all$residuals/Sflower_all)

#predicting kinkajou residence time based on study night and flower count

residence_regression_all <- lm(data = combined_data, formula = Minutes_Per_Hour ~ Study_night + Count)
summary(residence_regression_all)

#res_reg_all<-lm(data = combined_data, formula = Minutes_Per_Hour ~ Study_night)
#summary(res_reg_all)
# plot(x = combined_data$Study_night, y = combined_data$Minutes_Per_Hour)
# lines(x = combined_data$Study_night, y = residence_regression_all$coefficients["(Intercept)"] + residence_regression_all$coefficients["Study_night"] * combined_data$Study_night)
# 
# plot(x = combined_data$Count, y = combined_data$Minutes_Per_Hour)
# lines(x = combined_data$Count, y = residence_regression_all$coefficients["(Intercept)"] + residence_regression_all$coefficients["Count"] * combined_data$Count)

Sres_all <- sqrt(sum(residence_regression_all$residuals ^ 2)/(length(residence_regression_all$residuals) - 2))
plot(residence_regression_all$residuals,
     xlab = "Index",
     ylab = "Residuals",
     main = "Full Tree Residence Time Residuals")
qqnorm(residence_regression_all$residuals/Sres_all)
```

```{r flower count regression analysis by region}

# using filter function to create new flower data subsets
flowers_R1 <- filter(flower_counts_region, Region == 1) %>% as.data.frame() %>% select(!geom)
flowers_R2 <- filter(flower_counts_region, Region == 2) %>% as.data.frame() %>% select(!geom)
flowers_R3 <- filter(flower_counts_region, Region == 3) %>% as.data.frame() %>% select(!geom)

#predicting flower counts within the regions based on study night

# linear regression flower count region 1
flowers_regression_R1 <- lm(data = flowers_R1, formula = Count ~ Study_night)
summary(flowers_regression_R1)

plot(y = predict(flowers_regression_R1, newdata = data.frame(Study_night = 0:max(flowers_R1$Study_night))), x = 0:max(flowers_R1$Study_night), type = "l",
     xlab = "Study Night",
     ylab = "Flower Count",
     main = "Region 1 Flower Count Regression")
points(x = flowers_R1$Study_night, y = flowers_R1$Count)

plot(x = flowers_R1$Study_night, y = flowers_regression_R1$residuals,
     xlab = "Study Night",
     ylab = "Residuals",
     main = "Region 1 Flower Count Residuals")

Sflower_region1 <- sqrt(sum(flowers_regression_R1$residuals ^ 2)/(length(flowers_regression_R1$residuals) - 2))
qqnorm(flowers_regression_R1$residuals/Sflower_region1)

# linear regression flower count region 2
flowers_regression_R2 <- lm(data = flowers_R2, formula = Count ~ Study_night)
summary(flowers_regression_R2)

plot(y = predict(flowers_regression_R2, newdata = data.frame(Study_night = 0:max(flowers_R2$Study_night))), x = 0:max(flowers_R2$Study_night), type = "l",
     xlab = "Study Night",
     ylab = "Flower Count",
     main = "Region 2 Flower Count Regression")
points(x = flowers_R2$Study_night, y = flowers_R2$Count)

plot(x = flowers_R2$Study_night, y = flowers_regression_R2$residuals,
     xlab = "Study Night",
     ylab = "Residuals",
     main = "Region 2 Flower Count Residuals")

Sflower_region2 <- sqrt(sum(flowers_regression_R2$residuals ^ 2)/(length(flowers_regression_R2$residuals) - 2))
qqnorm(flowers_regression_R2$residuals/Sflower_region2)

# linear regression flower count region 3
flowers_regression_R3 <- lm(data = flowers_R3, formula = Count ~ Study_night)
summary(flowers_regression_R3)

plot(y = predict(flowers_regression_R3, newdata = data.frame(Study_night = 0:max(flowers_R3$Study_night))), x = 0:max(flowers_R3$Study_night), type = "l",
     xlab = "Study Night",
     ylab = "Flower Count",
     main = "Region 3 Flower Count Regression")
points(x = flowers_R3$Study_night, y = flowers_R3$Count)

plot(x = flowers_R3$Study_night, y = flowers_regression_R3$residuals,
     xlab = "Study Night",
     ylab = "Residuals",
     main = "Region 3 Flower Count Residuals")

Sflower_region3 <- sqrt(sum(flowers_regression_R3$residuals ^ 2)/(length(flowers_regression_R3$residuals) - 2))
qqnorm(flowers_regression_R3$residuals/Sflower_region3)
```

```{r residence regression analysis by region}

#predicting kinkajou residence time based on study night and flower count

# combining flower count and residence data containing the regions
combined_data_region <- merge(flower_counts_region, kinkajou_counts_region, by = c("Study_night","Region") , all = TRUE)

R1_data <- filter(combined_data_region, Region == 1)

#linear regression residence time region 1

#R1_Predictions <- predict(residence_region1_regression, newdata = select(R1_data, Count, Study_night))
#plot(x = filter(combined_data_region, Region == 1)$Study_night, y = filter(combined_data_region, Region == 1)$Minutes_Per_Hour, xlab = "Study Night", ylab = "Residence Time", main = "Residence Region 1 Regression") #plot actual residences times by night

#lines(x = filter(combined_data_region, Region == 1)$Study_night, y = R1_Predictions

# residence_region1_regression <- lm(data = R1_data, formula = Minutes_Per_Hour ~ Study_night + Count)
# summary(residence_region1_regression)
# 
# Sres_region1 <- sqrt(sum(residence_region1_regression$residuals ^ 2)/(length(residence_region1_regression$residuals) - 2))
# plot(residence_region1_regression$residuals, xlab = "Index", ylab = "Residuals", main = "Residence Region 1 Regression Residuals")
# qqnorm(residence_region1_regression$residuals/Sres_region1)
# 
# 
# 
# #linear regression residence time region 2
# residence_region2_regression <- lm(data = filter(combined_data_region, Region == 2), formula = Minutes_Per_Hour ~ Study_night + Count)
# summary(residence_region2_regression)
# 
# Sres_region2 <- sqrt(sum(residence_region2_regression$residuals ^ 2)/(length(residence_region2_regression$residuals) - 2))
# plot(residence_region2_regression$residuals, xlab = "Index", ylab = "Residuals", main = "Residence Region 2 Regression Residuals")
# qqnorm(residence_region2_regression$residuals/Sres_region2)
# 
# #linear regression residence time region 3
# residence_region3_regression <- lm(data = filter(combined_data_region, Region == 3), formula = Minutes_Per_Hour ~ Study_night + Count)
# summary(residence_region3_regression)
# 
# Sres_region3 <- sqrt(sum(residence_region3_regression$residuals ^ 2)/(length(residence_region3_regression$residuals) - 2))
# plot(residence_region3_regression$residuals, xlab = "Index", ylab = "Residuals", main = "Residence Region 3 Regression Residuals")
# qqnorm(residence_region3_regression$residuals/Sres_region3)

# combining region regressions into 1 regression
combined_data_region <- mutate(combined_data_region, Total_Flowers = combined_data$Count[pmatch(Study_night, combined_data$Study_night, duplicates.ok = TRUE)])
residence_regression <- lm(data = combined_data_region, formula = Minutes_Per_Hour ~ Count + Region + Total_Flowers)
summary(residence_regression)
# We want to visualize the affect of region and regional flower count, after accounting for the effect of total flower count. TO do this, we substract the modeled effect of total flower count from all of our residence times
combined_data_region = mutate(combined_data_region, Adjusted_Residence = Minutes_Per_Hour - Total_Flowers * residence_regression$coefficients["Total_Flowers"])

ggplot(combined_data_region, aes(x = Count, y = Adjusted_Residence, color = Region)) +
  geom_point() +
  geom_abline(slope = residence_regression$coefficients["Count"], 
            intercept = residence_regression$coefficients["(Intercept)"], colour="firebrick1") + #Region 1
  geom_abline(slope = residence_regression$coefficients["Count"], 
            intercept = (residence_regression$coefficients["(Intercept)"] + residence_regression$coefficients["Region2"]), colour="forestgreen") + #Region 2
  geom_abline(slope = residence_regression$coefficients["Count"], 
            intercept = (residence_regression$coefficients["(Intercept)"] + residence_regression$coefficients["Region3"]), colour="dodgerblue3") + #Region 3
  theme_classic()+
  labs( x = "Regional Flower Count", 
        y = "Adjusted Residence Time")

```
